Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-10-23 paper/main.tex
fix indices

We use principles from discrete differential geometry~\cite{bobenko2008discrete} to describe the
stiffness of the bacterial rod~\cite{Amir2014_2}.
As explained above, MReB will counteract the bending of the rod, resulting in a stiffening force.
We model the bacterium as a bending beam.
% In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988},
% but for sake of simplicity we chose to model the bacterium as a bending beam.
% \textbf{[", thus utilizing the simplest solution which can still produce the desired effects." Es ist unklar was die "desired effects" sind. ]}\\
Using our discretization $\vec{x}_i$, we connect the vertices by springs of length $l_i$.
% which can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
% springs of lengths $l_i$.
% \textbf{[Das klingt nach einer Wiederholung von Spatial Representation. Du erklärst ziemlich umständlich. Konzentriere dich auf das wesentliche und überlege, was du eigentlich sagen möchtest.]}\\
We use the notation $\vec{w}i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
between two vertices if given by

\begin{equation}
    \vec{F}_{i,\text{springs}} =
        -\gamma\left(1 - \frac{l}{\left|\vec{w}i\right|}\right) \vec{w}i
        + \gamma\left(1 - \frac{l}{\left|\vec{w}{i+1}\right|}\right) \vec{w}{i+1}.
\end{equation}

The value of the spring constants determines the elasticity with respect to elongation of the cell.
In situations where this aspect is not relevant, they merely serve as a auxiliary parameters to
maintain the length of the rod.
In addition to springs between individual vertices $\vec{x}_i$, we assume that each angle at a
vertex between edges is subject to a force indiced by curvature.
% We assume that we can model the mechanical properties of the bacterium as an elastic rod.
Given the angle $\alpha_i = \sphericalangle(\vec{w}{i-1},\vec{w}i)$ between adjacent edges,
the bending force is proportional to the curvature $\kappa_i$ at each vertex
$\kappa_i = 2\tan(\alpha_i/2)$.
The resulting force acts along the angle bisector which can be calculated from the edge vectors.
Figure~\ref{fig:mechanics-bacterium} shows an elongated rod represented by $6$ vertices.
The force acting on vertex $\vec{x}_i$ is given by

\begin{equation}
    \vec{F}_{i,\text{curvature}} = \eta\kappa_i
        \frac{\vec{w}i - \vec{w}{i+1}}{|\vec{w}i-\vec{w}{i+1}|}
\end{equation}

>>>>

We use principles from discrete differential geometry~\cite{bobenko2008discrete} to describe the
stiffness of the bacterial rod~\cite{Amir2014_2}.
As explained above, MReB will counteract the bending of the rod, resulting in a stiffening force.
We model the bacterium as a bending beam.
% In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988},
% but for sake of simplicity we chose to model the bacterium as a bending beam.
% \textbf{[", thus utilizing the simplest solution which can still produce the desired effects." Es ist unklar was die "desired effects" sind. ]}\\
Using our discretization $\vec{x}_i$, we connect the vertices by springs of length $l_i$.
% which can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
% springs of lengths $l_i$.
% \textbf{[Das klingt nach einer Wiederholung von Spatial Representation. Du erklärst ziemlich umständlich. Konzentriere dich auf das wesentliche und überlege, was du eigentlich sagen möchtest.]}\\
We use the notation $\vec{w}_i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
between two vertices if given by

\begin{equation}
    \vec{F}_{i,\text{springs}} =
        -\gamma\left(1 - \frac{l}{\left|\vec{w}_i\right|}\right) \vec{w}_i
        + \gamma\left(1 - \frac{l}{\left|\vec{w}_{i+1}\right|}\right) \vec{w}_{i+1}.
\end{equation}

The value of the spring constants determines the elasticity with respect to elongation of the cell.
In situations where this aspect is not relevant, they merely serve as a auxiliary parameters to
maintain the length of the rod.
In addition to springs between individual vertices $\vec{x}_i$, we assume that each angle at a
vertex between edges is subject to a force indiced by curvature.
% We assume that we can model the mechanical properties of the bacterium as an elastic rod.
Given the angle $\alpha_i = \sphericalangle(\vec{w}_{i-1},\vec{w}_i)$ between adjacent edges,
the bending force is proportional to the curvature $\kappa_i$ at each vertex
$\kappa_i = 2\tan(\alpha_i/2)$.
The resulting force acts along the angle bisector which can be calculated from the edge vectors.
Figure~\ref{fig:mechanics-bacterium} shows an elongated rod represented by $6$ vertices.
The force acting on vertex $\vec{x}_i$ is given by

\begin{equation}
    \vec{F}_{i,\text{curvature}} = \eta\kappa_i
        \frac{\vec{w}_i - \vec{w}_{i+1}}{|\vec{w}_i-\vec{w}_{i+1}|}
\end{equation}
